---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## Read in some data

Read in the litters dataset.

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
can load janitor package in the setup
but we only use one function, we don't need to do that


## Take a look at the data

Pirnting in the console 
```{r}
litters_df
```
<dbl> perticular kind of number
numerical variable

```{r}
head(litters_df)
```

```{r}
tail(litters_df)
```

```{r}
skimr::skim(litters_df)
```

view(litters_df)
open another window
better use in the console


##  Options to read_csv

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE, na = c("","NA",".","99999"))
```

check out ?read_csv() for more information

## Other file formats

Read in an excel file.

```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```
if multiple sheet - specify which one

Read in a SAS file 

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

## Comparison with Base R

what about `read.csv`...?

```{r}
litters_base = read.csv("data/FAS_litters.csv")
litters_readr = read_csv("data/FAS_litters.csv")

litters_base
litters_readr
```

